\chapter{The Siege}

\begin{center}
\begin{tcolorbox}[width=35em]

	As you cross the next hill, you see the stone walls of the town ahead, and the smoke coming up from the hundreds of chimneys within.
	Climbing higher, you can see ladders poking at the town's tall walls.
	A man runs along the edge, pushing them aside.

	You climb higher, and see the town surrounded by energetic, deformed, creatures of all sizes.
	The nura have multiplied more than you thought.
	They have laid siege to the entire town.

\end{tcolorbox}
\end{center}

\section{The Plan}

\begin{multicols}{2}

The party probably thought they were out of the water, but you can continue the adventure with the siege.

\subsection{The Scout}

\begin{boxtext}

	You see a man wandering towards you, wearing thick, black, leather armour.
	He seems to be a member of the Night Guard.
	As he stumbles closer, you notice the trickle of blood leaking a river behind him.

\end{boxtext}

\begin{exampletext}

	Laith and his men were charged with escaping the city, and delivering a message to `Baron Quenn' (as he likes to call himself), otherwise known as the `Black Alchemist'.
	The message would pardon him of all wrongdoings, and recognize his current lands as an official domain, if only he approaches the town to save it.

	Unfortunately, Laith and his men were then attacked by dozens of goblins.
	All died, except a heavily wounded Laith, who is bleeding heavily.

\end{exampletext}

The moment Laith approaches, he explains his mission, and gives the party directions.

\begin{speechtext}

	You must walk East until the Sun sets, find the place in the forest where the trees neither grow, nor die, then scout for the Black Alchemist who lives there, and hand him this message.
	The town needs him.

\end{speechtext}

\paragraph{At this point,}
the party spot a goblin sniffing along the stream of blood Laith left behind, far in the distance.
He shrieks in delight, then runs back.

\paragraph{If the party leave Laith behind immediately,}
they can run to safety, but will still hear Laith being torn apart.

If they wait to speak with him, a horde of goblins and ogres run fast behind them.

\ogre

\goblin

\end{multicols}
