\chapter{Timeline}

\section{History}

\begin{multicols}{2}

Ten years ago, the gnomes of Badger's Warren made a magical portal to the Realm of Bright Rocks in order to start making excursions there, and steal the precious flowers which produce such excellent magical items.

The deep creatures from the Realm of Darkness and Fire had their own portal in the Realm of Bright Rocks.
They don't usually journey to that domain much, as there is no food, and they do not like the Sun.
However, the goblins and ogres of that deep realm soon found the entrance to the gnomish warren.
They entered, grabbed the little gnomes, then ate them.
As they made their way up the warren, filling their stomachs, they started taking captives, and then used their Saurecanta magic to turn the gnomes into more nura.

Many nura died as they fell victim to the various traps the gnomes had laid, but there were always more nura coming to fight and die.
The nura-gnomes (`goblins') and the old nura together remade the warren into a war-ready stronghold.

At this point, a dragon from the Realm of Bright Rocks wandered into the portal.
The dragon -- `Makil' -- cannot fit through the small door to the treasure room it found, so he decided to just sleep there, waiting for someone small enough to help him.
The nura have had to leave the gold and precious magical items of this room well alone.

The next day, the nura went above-ground to raid various human villages close to the warren, and take those humans back to become food, or to be transformed into ogres.
The day after that, they used those ogres to bring back more people, completing their army.

On the third day, they laid siege to the local town.
They cannot get beyond the tall, stone walls, but that raiding party has plenty of food brought from the villages nearby.

\end{multicols}

\section{Plans}

\begin{multicols}{2}

\subsection{The Escape}
\label{escape}

\begin{boxtext}

	You awake in darkness with ropes binding your hands behind your back.
	From the shifting and murmuring sounds around, you know others rest in the same room.

\end{boxtext}

Here the PCs awaken to their hopeless situation.
Have everyone roll Dexterity + Larceny against TN 11 to escape.

\begin{boxtext}

	Heavy footsteps pad down the hall, you hear the door's bar being lifted, and a little goblinoid face peeps in, with a massive sack over her shoulder.
	Behind her, an ogre stoops to the height of a man to avoid the low ceiling while she unties the sack.
	The contents are chucked into the room -- apples, pies, raw potatoes, and lettuce, all in a pile on the filthy floor.
	James shouts to the ogre `Alf -- don't do this! Help me!', but the ogre only stares back at him before the little goblin shuts the door.

\end{boxtext}

James bleats for a moment before he crawls over and begins eating the food.
He knows what will happen to him, but he cannot stand to wait, starving in a cell, any more.

Resolve the scene, then check out the cell's description on page \pageref{entrycell}.

\paragraph{If the party kill James immediately}
then this is a good time to introduce the Sneak Attack rules.
James will be very vulnerable to attack in this state.
Remember also that if the party get a Vitals Shot in, they can deal Damage, rather than just inflicting Fatigue.

\goblin[\npc{\F}{Blara the Goblin}]

\ogre[\npc{\M}{Alf the Ogre}]

\subsection{The Raiding Party Return}
\label{raidingParty}

A full 20 ogres and 20 goblins have departed to grab prisoners from nearby villages.
15 of the ogres have returned with 10 prisoners.
5 of the goblins have come back -- many are still scouting outside.

They enter through the main door and make their way down slowly.
Every scene, they travel they travel through another room, until they reach the lift at the mid-floor position, then take the prisoners to room 1, where the PCs awakened.

These ogres have shed human blood, and have already forgotten about most of their original lives.
As they return and talk about how much they ate, the other ogres begin to accept their new lives, and accept that they must kill humans to live.
From this point onwards, ogres will not hesitate to enter battle.

\ogre[\npc{\T\N}{10 ogres}]

\ogre[\npc{\T\N}{5 ogres handling prisoners}]

\goblin[\npc{\T\N}{5 goblins}]

The ten prisoners do not have the strength to fight, but will flee if pushed.

\subsection{Captured}

If the party ever lose a fight, do not push this until each lie dead.
Instead, when it becomes obvious they cannot win, have the nura draw back, mock the party, and then tell them to drop their weapons so a group of ogres can escort them down to their cells.

If any of the party have died, more prisoners come in soon, so another character can be rolled and added to the story from this new group once someone in the party spends the necessary Story Points.

\end{multicols}
