\chapter{The Tour}

\begin{multicols}{2}
\setcounter{age}{0}

\begin{exampletext}
\noindent
Most gnomish warrens have little doors, but every rule has an exception.
Take this door for example -- big enough to fit a horse in.
It seems like they're just asking for some giant arachnid to rush in once the door opens and eat everyone inside, but in the \gls{warren} the architecture itself helps to defend the place.

Come on down, and I'll show you.

You see?
It's a spiral staircase!
Anything crawling down found gnomes at the bottom, with arrows an alchemy.
You wouldn't expect a little warren to have such a massive chamber down here.
You can hear every footstep echoing.
\forwards{stairs}

\smolMapPic[\large]{Dyson_Logos/upper}{
  \ref{tr:stairs}/31/59,
  \ref{tr:feast}/63/30,
}

The gnomes have gone already, except for a couple hiding in that room over there, with the door.
Goblins and ogres ate the rest I'm afraid.
See those guys playing dice?
Well they're much bigger close up, and not exactly `guys'.

`\textit{Hi guys!}'

(say `hi', and look friendly)

They're okay right now, they've already eaten.
They're under strict instructions to stop people getting out, but we're coming \textit{in}, and ogres often understand things too literally.

The rest of the journey we'll need some light.
Could you hold the torch here?
Keep it steady while I make some sparks.

Good, now how about you hold onto it?
Keep it to the side of your head -- not in front -- so it won't blind you.
We'll have plenty of room on the stairs.

Down the far door, we can't stop at the feasting table I'm afraid, unless you want to eat some human meat.
But look at the little tables!
The feasts here must have looked really cute before all this mess!

Listen!
That sounds like they're returning, and they're singing.
That means they have a good catch.
They don't have nearly enough food here, so they've had to go out to the local \glspl{village} and hunt for some human.
It's not their favourite, but it'll have to do them until they can find their way back home.
\forwards{feast}

Let's just stand back and let them pass by, down the stairs.
Did you see that one goblin just mouthing the words?
I wonder what his story is?

Yip, there they go, five prisoners in hand, all tied up, like lobsters with bound claws.
It's always funny seeing the ogres crawl their way through these little tunnels.
I hear if you get between two ogres in a small tunnel, the torch goes out from the lack of fresh air!

Okay, let's follow the sound of singing down.
This last door has a good, solid lock.
Shame it never helped the gnomes.

We'll need to take the big left-hand double-doors here.
The doors only go one way (so people don't bump into each other), so this way leads \emph{into} the warren, while the other only leads out.
\forwards{doors}

I'll open them, since you have the torch, and there they all are, torches in hand, bright as a hearthfire.
\forwards{lift}

Notice that a couple of the ogres and a goblin went into the next room, while the rest just wait for them?
We'll have to wait out turn for the room too.
Inside, an alchemically enchanted floor goes down the way, but slowly, or it can lift you up, if you're not too heavy.
Two ogres and a goblin can go down at a time, but not more, or the floor won't have any `lift' -- it'll just fall, and the ogres would die.
That's why the prisoners have to remain up here with one ogre, and all these goblins.
The goblin at the bottom has to use this `lift' to rise up to the top again, where they'll poke and prod at the humans to herd them into the little `lift-room'.

This is taking a while, but no price is too high for safety.

\smolMapPic[\large]{Dyson_Logos/lower}{
  \ref{tr:doors}/50/94,
  \ref{tr:lift}/32/77,
  \ref{tr:guard}/46/68,
  \ref{tr:workshop}/70/59,
  \ref{tr:babies}/43/48,
  \ref{tr:druid}/48/28,
  \ref{tr:dragon}/09/46,
}

Wait a minute -- I haven't thought this through.
Who's going to let \emph{us} in?
They won't come back up for us; their bleached little heads only have room for food right now.

Okay, down this way -- through the door.

Hurry up, and remember I'm blind until you bring the torch in.
It's pitch dark down here where I am.

Wait, see that hairy, maggot-white lump in front of the door?
That ogre must have fallen asleep.
It's okay -- he can't understand us.
Or `she'?
I can't usually tell, even by the chest.
They're all very bulbous with random proportions.
\forwards{guard}

Just don't disturb the ogre, or the prisoners in the door behind and we'll be fine.
Just need to squeeze past, gently\ldots

Down we go, keep up!
But I have a little light down here at least.
See all the fireflies?
The gnomes bred them in the next-room-down, so they could light the area.
Beautiful, eh?
Might be more beautiful without all the festering blood and tiny skulls.

Notice that the air's not too bad here?
Through those two arches, the gnomes built a workshop, with a forge.
Behind it, a chimney.
It started as a tunnel down, and still had a ladder, built into the stoneworks.
\forwards{workshop}

Anyway -- down we go, and quickly!
The once-tiny-slugs here have grown massive.
The goblin druids decided to do their usual tricks, making them into gargantuan monsters and ride them into battle.
Unfortunately, they couldn't adjust the speed properly, so they're just staying down here, spraying acid at anyone who gets close, while giant fireflies eat into them.

Magic, eh?
Never ends well.
Come on, I can hear the children screaming down the hall.

Here we are -- the gnomes locked their children in here (with plenty of food) before the goblinoid horde at them.
Now the cupboard has a dozen orphans, and the goblins have just learnt to ignore the screams.
\forwards{babies}

Down here, more double-doors, don't step through!
We need to wait for the floor to lift up some goblins.
They'll come back up once the prisoners are locked inside.
Here they are now.

`Scuse me!

Just squeeze back, ogres don't have any manners.
Goblins do, but they don't use them outside of courtships.

Okay, in we get.
All in?
Now what was the password?
I'd better remember fast, or the torch is going to swallow all the air in this little room, and we'll end up with hypoxia from the\ldots
sleep!

`\textit{Time to sleep!}'
(that's the word which makes the floor go gently down to the bottom)

Down we go!

Out we go.

Ha!
See that little guy with the scrunched-up face?
He's trying to decipher an alchemical scroll the gnomes left behind.
It's just like the ones used to make a gateway down to the goblin realms before.
\forwards{druid}

If he can just get the riddle written there, the doorway will open.
Of course, he can't read or speak any of \gls{fenestra}'s languages, and therein lies the tragedy of this whole thing.
If the goblins could only speak, they might have talked this whole thing out.

Aw, listen to him!
He's going through every sound he's ever heard the gnomes make, trying to get it to open the gateway to the \gls{deep}.

Okay, finally, let's meet the prisoners, actually the \glspl{pc}.

\ldots hang on.
No hang on.
The gnomish language doesn't have very many sounds, and the riddle on that scroll has a one-syllable answer.
He might actually get this by change.
He could actually open a gateway to somewhere deep below, then\ldots

Oh crap!
That's Sunlight coming through the other side of the gateway.
Which probably means, yep, it's a dragon, in he comes, bold as brass.
There the goblins go, looks like they're going to be dinner this time ar\ldots nope he's spat them out.
Let's just stay in here while he does his thing.

Up the narrow staircase he goes, straight to the treasure room.
Dragons can actually \emph{smell} different kinds of metal, especially when it's enchanted.

If he can barely squeeze up the stairs, how will he fit through the door?
That's the sound of a little broken lock, so he's got through the door, but listen\ldots
\forwards{dragon}

He can't fit through.
The dragon can't fit into the treasure room!
And how will he get back to his desert home without another gateway-summoning scroll?
Well that's his problem now.

So now you know the layout and what everyone's up to.
Give the schematics a proper read-through, and you'll be ready to open that last door, and meet the \glspl{pc}.

\end{exampletext}
\end{multicols}
