
\needspace{10\baselineskip}

\subsection{Mid Levels}
\iftoggle{intro}{
  The bulk of the horde lie in these rooms, napping.
  The \glspl{pc} may steal weapons or powerful \glspl{talisman}, and may free prisoners who can help them escape.
  Or they might tread carefully, and reach the top of the \gls{warren} silently.
}{}

\mapentry[nursery]{Nursery}

\begin{boxtext}
  Round the corner, a goblin squats back against a short, brass door.
  His eyes widen, as if making a quick plan.
\end{boxtext}

\begin{exampletext}
  When the horde arrived, a gnome locked the children in here with enough food for a few days.
  Earlier, \gls{kalama} the gnome cast a spell to look like a goblin, and sneaked in find any survivors.
  However, the spell drains his energy, making him confused and famished.
  He has found the children, and spotted the key next door, but cannot get to it.
\end{exampletext}

\spyGnome{2}

\paragraph{Opening the door}
requires the key in room \vref{nursery_key}.
Characters with a lock-picking set might also try to pick it with a \roll{Dexterity}{Larceny} roll (\tn[11]).
Breaking it open by force uses \roll{Strength}{Larceny} at \tn[20], so not even the ogres could manage.

If the \glspl{pc} enter the room, they may think these little gnomes are little goblins and kill them.
If any of them try to do so, have them roll an \roll{Intelligence}{Medicine} (\tn[8]) check to realize their mistake.
Of course, \gls{kalama} can identify the children immediately.

\begin{boxtext}
  The sound of crying emanates from the door as you swing it open.
  In this cramped room, a dozen infants with fat little noses lie in a crib with hay, looking up at you in terror.
  The tiny room stinks of shit.
\end{boxtext}

\paragraph{If the \glspl{pc} take the children,}
\gls{kalama} agrees to share the word to order the lift to go up to the top.
However, they will also scream at any disturbance.
Keeping the children quiet requires an \roll{Intelligence}{Empathy} roll (\tn[10]).

The lift can only ascend with a combined \gls{weight} of 17 or less, so the troupe will not be able to all go up at once, with the children, and \gls{kalama}.

\swarm{Gnomish Children}{5}{-2}{-3}{-5}

There are five children in total, and they each have \pgls{weight} between 1 and 2, but can generally be treated as a single `\gls{swarm}'.

None can walk, so the troupe must carry an additional \gls{weight} of~5, distributed as they please.
To further complicate matters, the gnome-children must go \emph{somewhere}, such as a bag.
Characters might also construct a baby-wrap out of clothes with an \roll{Intelligence}{Crafts} roll (\tn[10]).
Rolling a tie indicates the wrap does not work, while rolling a failure means the wrap works until the character tries to run\ldots

\paragraph{If the \glspl{pc} refuse to help the children,}
\gls{kalama} will wait until the horde attacks them, and then try to hunt for any remaining \glspl{talisman} in room \vref{grandLibrary}.
\nameref{grandLibrary} held many, and he knows where they were, and how they work.
If he finds them, he'll engineer a plan to save the children.

Exactly how this plays out depends on where the horde moves, whether or not the \glspl{pc} provide a distraction, and any other chaos the players throw into the mix.

\iftoggle{hardcore}{}{
  \warrenMapMiddle
}

\mapentry[slugHall]{Slug Hall}

\begin{boxtext}
  The doorway reveals a massive hallway of sparkling, floating, gently-buzzing, lights.%
  \footnote{Gnomes grew mushrooms throughout this room in order to grow slugs, which fed fireflies.
  While torches work better than fireflies, having omnipresent little lights wandering around the \gls{warren} made sure that people could get about easier.}
  Over to the left, a massive staircase leads up into the darkness.
  And ahead, the little lights gently illuminate giant slugs, feasting on corpses and torn-up books.

  The moment you enter, the slugs' eye-stalks perk up, and they begin to slide off the corpses they were feasting on, and approach\ldots
\end{boxtext}

The moment \pgls{pc} enters the room, a morphed slug projectile-vomits acid at them.
Each round after that, another slug comes within range, then remains where they are, blocking the return.%
\footnote{Once the goblin druids arrived, they cast life-engorging spells to grow the slugs to monstrous proportions, in order to ensure that prisoners don't escape past this point.}

\paragraph{If the \glspl{pc} throw food anywhere,}
the morph-slugs head towards it rather than fight.%
\footnote{The goblins move through this room by throwing `food' (books, poop, or bones) towards the far side of the room.
Salt covers the stairs, preventing the slugs from moving upwards.}

\paragraph{If the \glspl{pc} want to investigate the corpses,}
\label{nursery_key}
they find a dead gnome with \pgls{deep} Scroll, and a discarded, half-eaten cloak, with a key to room \ref{nursery} in the pocket.%
\footnote{See \autopageref{saving_the_children} for how the key arrived here.}
\iftoggle{intro}{%
  As with the others, the \Gls{deep} Scroll requires a riddle to be answered, and vanishes once used.
}{}

\morphslug[\npc{\T[9]\A\R}{20 Morph Slugs}]

\paragraph{While moving up the stairs}
the \glspl{pc} see thick dunes of salt, laid across the lower steps to stop the giant slugs moving up.

Throwing a fistfull of salt thrown at one inflicts $1D6$~\glspl{ep} and makes it \gls{retreat}.
Scattering the salt from high up the stairs gives all nearby slugs \pgls{ep}, while a direct fistful of salt inflicts \dmg{0}~\glspl{ep}; in either case they \gls{retreat} immediately.

Removing the salt from the stairs allows the slugs to journey up, slowly, in their own time, probably within \pgls{interval}.
Once that happens, a battle begins; half the \glspl{ogre} above die, and all of the slugs die.

\mapentry[greatHall]{The Great Hallway}

\begin{boxtext}
  At the
  \iftoggle{hardcore}{bottom}{top}
  of the stairs, this massive chamber lies empty, except for the fireflies darting about, and piles of long bones on the filthy floor.
  To the right, there two short tunnels reverberate with assorted snoring sounds.
  Ahead of you, two grand staircases lead up into a dim but unwavering light.
  Below you, the entire floor is sticky and greasy.%
  \footnote{The hungry horde have torn apart live humans and cows on this floor, leaving a sticky stain, and well-licked bones all over the floor.}
\end{boxtext}

If the \glspl{pc} have kept quiet enough, they find the goblin horde napping in the dark, with a contented chorus of snores (from rooms \ref{workshop} and \ref{grandLibrary}).

\iftoggle{hardcore}{}{
  \paragraph{If the \glspl{pc} have come from room \ref{slugHall},}
  they will not see the staircase on their left immediately, but \emph{will} see it after doing literally anything (fighting, searching, et~c.).
}

\paragraph{If the \glspl{pc} tarry or talk,}
have them roll \roll{Intelligence}{Stealth} (\tn[8]), to get across the sticky floor without squelching too much.%
\iftoggle{intro}{%
  They make only one \gls{natural}, which counts for the whole group.
  Each Margin on the roll allows them an additional round before the horde wakes.
}{}

\iftoggle{hardcore}{
  If a single \emph{player} clearly gives in-character advice to another, the horde awaken.
}{}

\boxPair[t]{
  \label{workshopGoblins}
  \goblin[\npc{\T[\arabic{enemyNo}]\N}{\arabic{enemyNo} Goblins}]

  \addtocounter{enemyNo}{-2}

  \ogre[\npc{\T[\arabic{enemyNo}]\N}{\arabic{enemyNo} Ogres}]
}{
  \pic{Decky/armoury}
}

\mapentry[workshop]{The Workshop}

\begin{boxtext}
  Some hidden candle lights a complete mess of tables, metallurgical equipment, weapons, unlit torches, snoring \glspl{ogre} and goblins.
  The place is so full, you can't make out how many lie here, but the snoring indicates more than you can see.
  On the benches, most of the equipment lies broken, but delicate gnomish hands once used these tables to polish gems, craft magical items, and forge digging equipment.
  On one table, you can see a pile of greatswords, some shortswords, and a couple of spears.
\end{boxtext}

\paragraph{If the \glspl{pc} have come up from room \ref{escapeShaft},}
the reaction depends entirely on how much noise they made while pushing the great forge aside.

\paragraph{If the \glspl{pc} attempt to take either a sword or a spear,}
each attempt requires a \roll{Dexterity}{Stealth} roll, \tn[7].
Failure will awaken the entire horde, while a tie allows them a good head start.

\paragraph{If the \glspl{pc} feel overwhelmed fighting against the horde,}
they may wish to \gls{retreat} from the fight and flee.
Have them roll \roll{Speed}{Athletics} at \tn[10] to see how far they get.

Find the goblins' stats \vpageref{workshopGoblins}.

\mapentry[grandLibrary]{The Grand Library}

\begin{boxtext}
  At the top of the stairs you find the ruins of a massive library, with little candles dotted about the room's edge.
  \Pgls{ogre} sleeps on a bed of torn-up books, and a pale, little hooded creature lies on a little bed, clutching something to its chest, even in its sleep.
\end{boxtext}

The sleeping druid holds one \lootTalisman\ \gls{talisman}.

\showTalisman

\goblincaster

\iftoggle{hardcore}{
  \ogre[\npc{\T[2]\N\M}{2 Docile Ogres}]
}{
  \ogre[\npc{\N\F}{Sleeping Ogre}]
}

\paragraph{To investigate quietly,}
roll \roll{Dexterity}{Stealth} at \tn[8].
\labyrinthScroll 
Success reveals a useable \spellName.
Failure awakens the horde.

\mapentry[windingStairs]{Winding Stairs}

\begin{boxtext}
  \Pgls{ogre} slouches in front of a door, licking a long bone with a bright-red tongue.
  The ong legs block the path completely.
\end{boxtext}

\paragraph{If the party have made a reasonable attempt at staying quiet,}
they can avoid alerting this ogre with a \roll{Wits}{Stealth} roll, \tn[9].
Whoever is at the front makes the roll.
If it's unclear who's at the front, the character with the highest \roll{Speed}{Athletics} is in the lead.
With a successful roll,
\iftoggle{hardcore}{
  the ogre is resting, and must take a round to gather what's left of his wits, but it will still wake if approached.
}{%
 the party find the ogre sleeping.
}

\ogre[\npc{\N\M}{Prisoner Guard}]

\mapentry[secondPrison]{Second Prison}

\begin{boxtext}
  Opening the door opens into the room silently.
  A crumpled human lies on the tiny bed.
\end{boxtext}

This person only pretends to sleep so the other --- hiding round the corner of the room --- can sneak-attack%
\exRef{core}{Core Rules}{sneakattack}
whoever comes to eat them.

\iftoggle{intro}{
  If the troupe don't make it obvious they're not goblins, they'll have to roll \roll{Wits}{Vigilance} to avoid being stabbed.
  The \gls{tn} equals 7 plus the hiding human's \roll{Dexterity}{Stealth}
  (Select two \glspl{statblock} from the \gls{npc} handouts).
}{}

\iftoggle{hardcore}{
  \paragraph{If the \glspl{pc} came looking for their allies,}
  one lives here, unwounded, but probably terrified.
}{
  \paragraph{If any of the \glspl{pc} have died,}
  introduce another \gls{pc} here from the pool.
}

