\documentclass[10pt,twoside]{book}

\usepackage{config/bind}
\usepackage{config/booklet}
\input{glossary.tex}

\begin{document}

\miniCover{Escape from the Goblin Horde}{
  \par
  \vspace{1\baselineskip}
  \begin{minipage}{.8\linewidth}
    \pic{Irina/1}
  \end{minipage}
}

{
  \sffamily
  \noindent%
  John and Justin (from 2Legit2Crit) joined Ross and Couch to try their luck in a goblin-infested warren.
  We had the rare treat of an artist --- Irina -- drawing the plans, banter, and nonsense as the game progressed.

  \Gls{bind} modules change bit by bit, so the current module won't match these scenes precisely.
  And any other game won't match these scenes at all, as players never do what one expects.
}

\clearpage

\pagestyle{minizine}
\parindent0pt
\renewcommand\bigLine{}

\sidepic{Irina/2}{\textit{``We're all going to die down here.}}
Players began with their \glspl{pc} tied up in the dark, with memories slowly resurfacing of being captured by goblins. 
A morose man explained that goblins have eaten the gnomes who made the warren, and all the other prisoners, and they're going to come back down for more soon.

\begin{multicols}{2}
  \pic{Irina/3}

  \columnbreak

  The new prisoners tried to figure out how long he'd been down here by the size of his appetite. 
  He was the only person down there not tied up, so being friendly seemed like a good idea.
\end{multicols}

\noindent%
\begin{minipage}{\linewidth}
\begin{multicols}{2}
  \pic{Irina/4}

  \columnbreak

  Finally, a goblin druid arrived with an ogre.
  The time had come for them to take \pgls{pc} away to be eaten.
  \par
  \null

  Luckily they'd prepared to ambush him, using a plank of wood from the ground. 

\end{multicols}
\end{minipage}

\begin{multicols}{2}
  \pic{Irina/5}

  \columnbreak

  \pic{Irina/6}
\end{multicols}

\needspace{4cm}
The druid attempted an offensive spell, but someone managed to tear her head off while another stabbed the ogre in the brains. 

\pic{Irina/7}

\begin{multicols}{2}
  And more luck!
  The goblin druid had a fresh human foot in her bag, so the gnoll got to gnawing on it.

  \columnbreak

  \pic{Irina/8}
\end{multicols}

\begin{multicols}{2}
  \pic{Irina/9}{}

  \columnbreak

  Out the door, another goblin and a giant rat immediately attacked the elf.
  The new crew slew the pair, but were still fumbling in the dark, with no way to co\"ordinate.
\end{multicols}

\vspace{-2em}
\begin{speechtext}
\pic{Irina/10}
\end{speechtext}

\begin{multicols}{2}
  And when they found the kitchen, whatever root-vegetables remained were mixed with half-cooked human. 

  \columnbreak

  \pic{Irina/11}
\end{multicols}

\vspace{-1em}
\needspace{8em}
\begin{multicols}{2}
  \pic{Irina/12}

  \columnbreak

  Finding an empty room that goes nowhere, the group discussed the purpose, until they figured out it was a magical lift.
   But being alone and scared, they decided to look about.
\end{multicols}

They found stairs leading up to dead goblins --- charred more than the humans in the kitchen.

\pic{Irina/13}

Further up, a dragon looked down.
It was trapped in the warren, and unable to reach through the tiny gnomish door to the treasure-room.
The dragon spoke into their minds, demanding that they find a magical portal-scroll so he could return back the way he came, with the treasure. 

\sidepic{Irina/14}{}
So everyone emptied the treasure room, leaving the lot to the dragon.
The dragon cast a spell to compel one of the characters to speak the words to open the scroll (`cabbage').
The character managed to resist the enchantment just long enough to barter with the dragon --- they would give him the lot, but wanted some of the ogres dead

\begin{multicols}{2}
  The player had to name every vegetable he could which was \emph{not} cabbage in order to hold off.
  The plan worked, and the dragon left through the portal only after leaving some charred ogre-corpses. 

  \columnbreak

  \pic{Irina/15}
\end{multicols}

\begin{multicols}{2}
  \pic{Irina/16}

  \columnbreak
  Soon after, more ogres came down the lift, and the group managed to dispatch them, then go up the lift.
\end{multicols}

\sidepic[6]{Irina/17}{\textit{``You  ever taste cursed swamp-fruit?''}}
The lift couldn't take the weight of everyone at once, so the first group went up, leaving the gnoll and elf together.
The gnoll explained how humans taste, and the elf failed to explain that he didn't want to know.

\clearpage

Once everyone reached the top, they found a locked cupboard full of screaming gnomish children.
Seeing the gnoll did not help.

\pic{Irina/18}

\sidepic[4]{Irina/19}{\textit{``Oh no\ldots I think I may actually have to do this''}}
One of the warren gnomes had survived with a spell which warped his body to look like a goblin.  Unfortunately, he also acquired the goblinoid hunger.  When he couldn't take any more, he slunk into the darkness to eat the gnoll's half-masticated human foot. 

\begin{multicols}{2}
  \pic{Irina/20}

  \columnbreak

  The next room finally allowed the crew some light, as it was full of fireflies\ldots and acid spitting slugs.
  \par\null\par
  Up the stairs, the crew found a massive hallway, full of cavities, emanating the sound of snoring goblins and ogres.
\end{multicols}

So they found a silent passage, leading up.

\begin{multicols}{2}

  Two flights up, an ogre sat, half-asleep, guarding the stairs.
  They managed to rush the ogre-guard, but did not notice it was guarding a prison, so I guess everyone in the cell was eaten soon after. 

  \columnbreak

  \pic{Irina/21}
\end{multicols}

\begin{multicols}{2}
  \pic{Irina/22}

  \columnbreak

  Upstairs the characters found the exit-door locked.
  Going the other way, they found the lift again, and beside it a the goblin-druid, taunteing them with a set of keys.
  Beside him, alchemically mutated dogs barked, ready to kill.
\end{multicols}

\sidepic[4]{Irina/23}{\tiny\textit{``nom-nyom-mlem''}}
The party fled back to the lift, rather than fighting while (mostly) unarmed, and (entirely) unarmoured.
By this point, the goblin horde below had heard them, and emerged at the top of the lift-shaft, looking down.
A single ogre stood guard at the top, with a massive sword, while the rest went down to eat them.

\needspace{8em}

\sidepic[3]{Irina/24}{\tiny\textit{``nyah-nyah''}}
The elf decided to `taunt the ogre', and eventually the ogre became enraged, and jumped to its own death.
The plan to `taunt the ogre' did not sound like a good one, so I set the \gls{tn} to 16, plus the ogre's \roll{Intelligence}{Tactics}.
The ogre's Tactics was `bugger-all', and its Intelligence was -5,%
\stepcounter{footnote}\footnotemark[\thefootnote]
so the final \gls{tn} was 11.

\sidepic[6]{Irina/25}{\tiny\textit{``sploot''}}
The falling Damage was 4 + the ogre's Strength (the bigger they are, the harder they fall) so the ogre went `splat'. 

The crew ascended with the elevator (still carrying the gnome-children), ran after the goblin with the keys, and found themselves in a massive fungal-garden, where semi-sentient acidic oozes stalked them.

\begin{multicols}{2}
  \pic{Irina/26}

  \columnbreak

  \pic{Irina/27}
\end{multicols}

The raiding party of ogres and goblins began chasing after them while they ran after the little goblin-druid who still had the keys to the exit-door.

\sidepic[6]{Irina/28}{\tiny\textit{``sploot''}}
Two of the crew grabbed the little goblin by the keys, and threw him into the acidic ooze.
The others redirected the acidic oozes to chase the goblin horde.

With a `click', the door opened, and they all felt Sunlight.

This was the first play-through where a group managed to save all of the gnome-children.

\footnotetext{All \glspl{npc} have unique stats, so ogre has a different Intelligence Bonus.
You wouldn't think that would ever come up in a game, but apparently it does.}

\pic{Irina/29}

\end{document}
